---
title: "Exercise1_Week5"
author: "Lucy Bryson"
format: html
editor: visual
---

The following exercises are a test in disguise. Can you think of any improvements to the following code?

Go through the exercises and answer them while fixing issues and improving on code workflow

Make a Rmarkdown (or Quarto) version of this document with your responses

Render the document in PDF and HTML formats

```{r}
rm(list=ls()) 

library(tidyverse)
```

## PROBLEM 1

Visualizing your data is important!

Summary statistics can be highly misleading, and simply plotting the data can reveal a lot more! Lets look at the Anscombe’s Quartet data. There are four different data sets.

Anscombe, F. J. (1973). "Graphs in Statistical Analysis". American Statistician. 27 (1): 17–21. doi:10.1080/00031305.1973.10478966. JSTOR 2682899.

```{r}
anscombe_quartet = readRDS("anscombe quartet.rds")
```

### Exploring the Data Set 

```{r, exploration}
str(anscombe_quartet)
```

#### What does the function str() do?

It provides a list of variables to console, providing slight insight into the data

#### Let's check some summary statistics:

```{r, summary stats}
anscombe_quartet %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x    = mean(x),
    mean_y    = mean(y),
    min_x     = min(x),
    min_y     = min(y),
    max_x     = max(x),
    max_y     = max(y),
    crrltn    = cor(x, y)
  )
```

#### What do the summary statistics tell us about the different datasets?

The summary statistics say the data sets are similar, with data set 4 having a higher minimum and maximum value

#### Let's plot the data with ggplot:

```{r, plotting the data}
library(ggplot2)

theme_set(theme_bw(base_size = 15) + theme(legend.position="bottom"))

aq_plot <-
ggplot(anscombe_quartet, aes(x=x,y=y)) +
  geom_point() + 
  geom_smooth(method = "lm",formula = "y ~ x") +
    theme(panel.grid.minor = element_blank()) +
   coord_cartesian(ylim = c(0, 20), xlim = c(0, 20)) +
  facet_wrap(~dataset)
```

#### Save as plot - PNG format

Width 5 inch, height = 5 inch for paper

```{r, save plot}
ggsave(filename ='anscombe_quartet.png', plot = aq_plot, width = 5, height = 5)
```

#### What do the plots tell us about the different datasets?

Describe the relationship between x and y in the different datasets.

-   X and Y have a postive relationship in both data set 1 and 3

-   There is a normal distribution (seems like a U-shaped curve) in data set 2

-   All variables in data set 4 seem to have the same X value with similar X values

#### Would linear regression be an appropriate statistical model to analyse the x-y relationship in each dataset?

-   For data set 1 and 3 yes

-   For data set 2, as it is normally distributed, I would suggest using an ANOVA instead

-   For data set 4, perhaps a t-test represented as a bar chart would be more effective analysis

#### What conclusions can you draw for the plots and summary statistics?

## PROBLEM 2

Load in the datasaurus dataset.

```{r, load in data}
datasaurus_dozen = readRDS("datasaurus_dozen.rds")
```

#### Explore the different datasets in datasaurus_dozen

```{r, explore the data}
str(datasaurus_dozen)

datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x    = mean(x),
    mean_y    = mean(y),
    min_x     = min(x),
    min_y     = min(y),
    max_x     = max(x),
    max_y     = max(y),
    crrltn    = cor(x, y)
  )
```

#### Calculate the correlations for x and y

```{r, calculate correlations}
	r1 <- lm(x ~ y, datasaurus_dozen)
summary(r1)

```

#### Plot their relationships including the line of best fit.

```{r, plotting including line of best fit}
	library('GGally')
p2 <- ggplot(datasaurus_dozen, aes(x = x, y = y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  facet_wrap(~dataset) +
  labs(title = "Scatter Plots with Line of Best Fit",
       x = "X Axis",
       y = "Y Axis") +
  theme_minimal()
```

#### Save the plot

```{r, save the plot}
ggsave(filename ='datasauras_dozen.png', plot = p2, width = 5, height = 5)
```
